---
title: "Machine Learning Project"
author: "Aysegul Sonmez"
date: "March 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background  

###Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

```{r}

library(knitr)
library(rpart.plot)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
library(caret)
library(rpart)
library(RColorBrewer)
library(gbm)
library(plyr)
library(rpart)

dt_training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
dt_testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
options <- names(dt_testing[,colSums(is.na(dt_testing)) == 0])[8:59]

```
# Only use features used in testing cases.

```{r}

dt_training <- dt_training[,c(options,"classe")]
dt_testing <- dt_testing[,c(options,"problem_id")]
dim(dt_training); dim(dt_testing);
```

##Partitioning the Dataset

###Following the recommendation in the course Practical Machine Learning, we will split our data into a training data set (60% of the total cases) and a testing data set (40% of the total cases; the latter should not be confused with the data in the pml-testing.csv file). This will allow us to estimate the out of sample error of our predictor.

```{r}
library(caret)
set.seed(54321);

inTrain <- createDataPartition(dt_training$classe, p=0.6, list=FALSE)
training <- dt_training[inTrain,]
testing <- dt_training[-inTrain,]

dim(training); dim(testing);

```
                
##Building the Decision Tree Model

###Using Decision Tree, we shouldn't expect the accuracy to be high. In fact, anything around 80% would be acceptable.


```{r,echo=TRUE}
set.seed(54321)
modFitDT <- rpart(classe ~ ., data = training, method="class", control = rpart.control(method = "cv", number = 10))
fancyRpartPlot(modFitDT)
```


##Predicting with the Decision Tree Model

```{r}
set.seed(54321)
prediction <- predict(modFitDT, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```



##Building the Random Forest Model

###Using random forest, the out of sample error should be small. The error will be estimated using the 40% testing sample. We should expect an error estimate of < 3%.

```{r,echo=TRUE}
set.seed(54321)
modFitRF <- randomForest(classe ~ ., data = training, method = "rf", importance = T, trControl = trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10))
plot(modFitRF)

```
##Predicting with the Random Forest Model

```{r}
prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```
###The random forest model performed very well in-sample, with about 99.3% Accuracy.
##Building the Boosting Model

```{r,echo=TRUE}
##install.packages("caret", dependencies = c("Depends","Suggests"))
##install.packages("lubridate")
library("caret")
library("lubridate")

modFitBoost <- train(classe ~ ., method = "gbm", data = training,
                     verbose = F,
                     trControl = trainControl(method = "cv", number = 10))

modFitBoost
plot(modFitBoost)

```
##Predicting with the Boosting Model

```{r}
prediction <- predict(modFitBoost, testing)
confusionMatrix(prediction, testing$classe)

```
##Predicting with the Testing Data (pml-testing.csv)

###Decision Tree Prediction
```{r}
predictionDT <- predict(modFitDT, dt_testing)
predictionDT
```
###Random Forest Prediction
```{r}
predictionRF <- predict(modFitRF, dt_testing)
predictionRF
```
###Boosting Prediction
```{r}
predictionBoost <- predict(modFitBoost, dt_testing)
predictionBoost
```


##Conclusion
###As can be seen from the confusion matrix the Random Forest model is very accurate, about 99%. Because of that we could expect nearly all of the submitted test cases to be correct. It turned out they were all correct.


## Now we fit our Random Forest model to the original testing raw data
```{r echo=TRUE}
prediction_originaltesting_RF <- predict(modFitRF, testing, type = "class")

